---
output:
  word_document: default
  html_document: default
---
# Model Validation
## Katherine Schumann
### Mod 3 Assignment 1


```{r Libraries}
library(tidyverse) #tidyverse set of packages and functions
library(tidymodels)
library(lubridate)
```

```{r}
bike_cleaned <- read_csv("C:/Users/kathe/OneDrive/Desktop/Ban 502/Module 3/bike_cleaned.csv", 
    col_types = cols(dteday = col_datetime(format = "%m/%d/%Y %H:%M")))
bike = bike_cleaned
```
```{r}
#bike = bike %>% mutate(dteday = mdy(dteday))#mdy is a lubridate package function
bike = bike%>% mutate(season = as_factor(season))#mdy is a lubridate package function
bike = bike%>% mutate(mnth = as_factor(mnth))#mdy is a lubridate package function
bike = bike%>% mutate(holiday = as_factor(holiday))#mdy is a lubridate package function
bike = bike%>% mutate(weekday = as_factor(weekday))#mdy is a lubridate package function
bike = bike%>% mutate(workingday = as_factor(workingday))#mdy is a lubridate package function
bike = bike%>% mutate(weathersit = as_factor(weathersit))#mdy is a lubridate package function

bike = bike %>% mutate(hr = as_factor(hr))
```

### Task 1

Split  
```{r}
set.seed(1234)
bike_split = initial_split(bike, prob = 0.80, strata = count)
train = training(bike_split)
test = testing(bike_split)
```

### Task 2
There are 13036 rows of data in the training set and 4343 rows of data in the test set. 

### Task 3

```{r}
bike_recipe = recipe(count ~ season + mnth + hr + holiday + weekday + temp + weathersit, train)

lm_model = #give the model type a name 
  linear_reg() %>% #specify that we are doing linear regression
  set_engine("lm") #specify the specify type of linear tool we want to use 

lm_wflow = 
  workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(bike_recipe)

lm_fit = fit(lm_wflow, train)
```

```{r}
summary(lm_fit$fit$fit$fit)
```

Adjusted R square is 0.6219. 


```{r}
predict_train = predict(lm_fit,train)
```

```{r Histogram of Predictions}
ggplot(predict_train, aes(.pred))+
  geom_boxplot()
```
The above distribution looks normal with a mean of just under two hundered. The only issue is that it goes into negitive numbers for the .pred, which is in probable so I would use common sense and know that there was not negitive bike rides if I were to present this data.


```{r}
lm_fit %>% predict(test) %>% bind_cols(test) %>% metrics(truth = count, estimate = .pred)
```

The value of the R squard went up slightly but is pretty similar, like in the other videos I would say that due to the smaller size in data there is a chance for the r squared value to fluxuate slightly. 

